### 2022-09-21 HW2/HW2.Rmd
start hw@

---
title: "HW 2: Energy Demand and CO2 Emissions"
author: "Hayden Atchley"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: lualatex
mainfont: Gentium Book Basic
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(tidyverse)
library(gridExtra)
library(kableExtra)
```

# 1

```{r}
vm1 <- read_csv("HW2/data/vm1.csv", n_max = 28) %>% 
	mutate(ITEM = ifelse(is.na(ITEM), lag(ITEM), ITEM))

co2gal <- c(gasoline = 8887, diesel = 10180)
```

Creating a new category combining all buses, single-unit trucks, and combination trucks gives the following:

```{r}
hdv1 <- vm1 %>% 
	mutate("ALL HDV" = BUSES + `SINGLE-UNIT TRUCKS/3` + `COMBINATION TRUCKS`)

vmlong <- hdv1 %>% 
	pivot_longer(c(-YEAR, -ITEM), names_to = "category") %>% 
	arrange(category, ITEM)
```

## 1.1
The average fuel economy is

```{r}
hdvfe <- hdv1 %>% 
	filter(ITEM == "Average miles traveled per  gallon of fuel consumed") %>% 
	select(YEAR, `ALL HDV`) %>% 
	transmute(Year = YEAR, "Fuel economy (mpg)" = `ALL HDV`)

hdvfe %>% 
	kbl(booktabs = TRUE, caption = "Average Fuel Economy of HDVs") %>% 
	kable_styling(latex_options = "HOLD_position")
```

## 1.2
The CO~2~ emissions per mile is given by $\frac{1}{miles/gal} \times \frac{CO_2}{gal} = \frac{CO_2}{mile}$. This gives

```{r}
hdvem <- hdvfe %>% 
	mutate("Fuel consumption (gpm)" = 1/`Fuel economy (mpg)`,
				 "Emissions (g CO$_2$/gal)" = co2gal["diesel"],
				 "Emissions (g CO$_2$/mile)" =
				 	`Fuel consumption (gpm)` * `Emissions (g CO$_2$/gal)`)

hdvem %>% 
	kbl(booktabs = TRUE, escape = FALSE, caption = "HDV Fuel Consumption") %>% 
	kable_styling(latex_options = "HOLD_position")
```

## 1.3
The total CO~2~ emissions for all LDVs and HDVs is

```{r}
allem <- vmlong %>% 
	filter(category %in% c("ALL LIGHT DUTY VEHICLES/2", "ALL HDV"),
				 ITEM %in% c("Fuel consumed   (thousand gallons)")) %>% 
	pivot_wider(names_from = ITEM) %>% 
	mutate("Emissions (g CO$_2$/gal)" = c(rep(co2gal[2], 2),
																				rep(co2gal[1], 2)),
				 category = ifelse(str_detect(category, "LIGHT"),
				 									"ALL LDV", category)) %>% 
	arrange(desc(category), desc(YEAR)) %>% 
	mutate("Emissions (g CO$_2$)" = 
				 	`Fuel consumed   (thousand gallons)` *
				 	`Emissions (g CO$_2$/gal)` * 1000,
				 "Emissions (M tons CO$_2$)" = `Emissions (g CO$_2$/gal)` / 1000000)

allem %>% 
	kbl(booktabs = TRUE, escape = FALSE,
			caption = "Emissions of LDVs and HDVs") %>% 
	kable_styling(latex_options = c("HOLD_position", "scale_down"))
```