### 2022-11-08 HW4/HW4.Rmd
start hw4

---
title: "HW #4: Analyze real-world emissions data"
author: Hayden Atchley
date: "`r Sys.Date()`"
mainfont: Gentium Book Basic
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

pacman::p_load(tidyverse, kableExtra, zfit, broom, ggthemes, DescTools)
```

```{r}
feat <- read_csv("data/feat.class.example.csv")
```

# 1

```{r}
#| em-density,
#| fig.cap="Emissions density by date.",
#| fig.pos="H"

feat1 <- feat %>% 
	replace_na(list(NO_g_kg = 0, NO2_g_kg = 0, HC_g_kg = 0, CO_g_kg = 0)) %>% 
	mutate(NOx_g_kg = NO_g_kg + NO2_g_kg) %>% 
	select(DATE, CO_g_kg, HC_g_kg, NOx_g_kg, location, VEH_ID) %>% 
	pivot_longer(-c(DATE, location, VEH_ID), names_to = "pollutant", values_to = "amount") %>% 
	mutate(pollutant = str_remove(pollutant, "_.+"))

feat1 %>% 
	ggplot() +
	geom_density(aes(x = amount, color = pollutant)) +
	facet_wrap(~DATE) +
	lims(x = c(NA,200),
			 y = c(NA,0.1)) +
	labs(x = "Emmission Rate (g / kg fuel)",
			 y = "kernel density",
			 color = "Pollutant")
```

## 1.1

```{r}
feat1 %>% 
	group_by(location, pollutant) %>% 
	summarise(max_emisions = max(amount), median_emissions = median(amount)) %>% 
	mutate("max / median" = max_emisions / median_emissions) %>% 
	kbl(booktabs = TRUE, digits = 2, linesep = "") %>% 
	kable_styling(latex_options = "HOLD_position")
```

## 1.2

Looking at Figure \@ref(fig:em-density), NOx appears to have quite a skewed distribution, though CO has a few extreme outliers.

# 2

```{r}
#| fig.height=3

feat2 <- feat1 %>% 
	select(VEH_ID, pollutant, amount) %>% 
	pivot_wider(names_from = pollutant, values_from = amount)

feat2 %>% 
	ggplot(aes(x = CO, y = HC)) +
	geom_point() +
	geom_smooth(method = 'lm') +
	theme_pander()
```

```{r}
#| fig.height=3

feat2 %>% 
	ggplot(aes(x = CO, y = NOx)) +
	geom_point() +
	geom_smooth(method = 'lm') +
	theme_pander()
```

```{r}
fit2 <- feat2 %>% 
	zlm(CO ~ NOx + HC) %>% 
	summary()

fit2 %>% 
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(0,2,2,2,3), escape = FALSE) %>% 
	footnote(paste0("$R^2$ = ", fit2$adj.r.squared %>% round(3)),
					 general_title = "") %>% 
	kable_styling(latex_options = "HOLD_position")
```

While it appears that vehicles with more CO emissions also have more NOx and HC emissions (both of these slopes/coefficients are positive), the $R^2$ value is quite low. There could be many other factors explaining the variance in emission rates.

# 3

# 4

```{r}
feat1 %>% 
	filter(pollutant == "CO") %>% 
	select(-pollutant, -VEH_ID) %>% 
	group_by(DATE, location) %>% 
	summarise(
		name = names(MeanCI(amount, conf.level = 0.9)),
		value = MeanCI(amount, conf.level = 0.9)) %>% 
	pivot_wider()
```

```{r}
feat1 %>% 
	filter(pollutant == "CO") %>% 
	select(-pollutant, -VEH_ID) %>% 
	ggplot(aes(y = DATE, x = amount, fill = location)) +
	geom_violin() +
	stat_summary(
		fun.data = function(x) MeanCI(x, conf.level = 0.9) %>% 
			`names<-`(c("y", "ymin", "ymax")),
		geom = "errorbar",
		width = 0.3) +
	stat_summary(fun = mean, geom = "point")
```